%\documentclass[a4paper, review, endfloat, doubleblind, authoryear]{elsarticle}
\documentclass[a4paper, review, endfloat, authoryear]{elsarticle}

% Packages (optional)
\usepackage{amsmath}  % For mathematical symbols
\usepackage{graphicx} % For including images
\usepackage{hyperref} % For clickable links

% Title, author, and affiliation
\title{Digital Transformation in the Shipping Industry: a Network-Based Systematic Review}

\author[1]{Andreas Pittas}
\ead{20200114@stu.uol.ac.cy}

\author[1]{Yannes Filippopoulos}
\ead{y.filippopoulos@uol.ac.cy}

\author[2]{Zoran Lajic}
\ead{zlajic@angelicoussisgroup.com}

\author[1]{Luca Ferrarini\corref{cor1}}
\ead{luca@uol.ac.cy}

\cortext[cor1]{Corresponding author}

\affiliation[1]{organization={Department of Information Technologies, University of Limassol},
	city={Limassol},
	country={Cyprus}}
\affiliation[2]{organization={Department of Energy Efficiency, Angelicoussis Group},
	city={Athens},
	country={Greece}}

\begin{document}  % Begin the document
	
	\begin{abstract}
		The shipping industry is undergoing a profound digital transformation, driven by advancements in automation, artificial intelligence, blockchain, and the Internet of Things (IoT). These technologies enhance operational efficiency, optimize supply chain management, and improve sustainability by reducing emissions and fuel consumption. However, navigating this digital revolution requires a structured understanding of emerging trends, challenges, and opportunities.
		A network-based systematic review serves as a crucial methodological approach for researchers, enabling them to synthesize existing knowledge, identify research gaps, and develop informed strategies to leverage digital transformation effectively.
		By critically analyzing co-citation and co-authorship networks, modeling topics over time, and performing trend analysis, we gain insights on the current status of digital transformation within the shipping industry, ultimately guiding industry stakeholders and researchers ....XXXXXXXXXXXXXx
		Our results show that .....XXXXXXXXXXXXXXXXXXXXXX
	\end{abstract}
	
	\begin{keyword}
		digital transformation \sep shipping industry \sep systematic literature review \sep complex networks
	\end{keyword}
	
	\maketitle  % Generates the title, author, and abstract section
	
	\section{Introduction}
	XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX
	
	\section{Literature Review}
	
	\section{Methodology}
	In this section we describe the methodology we followed for the data collection and analysis. Fig. [XXX] shows the overall methodology discussed in this section. Results and implications are discussed in further sections.
	
	\subsection{Keyword identification and data collection}
	We asked experts in the shipping industry to identify the most relevant keywords related to the industry itself and to digital technologies and digital transformation. Their analysis resulted in 35 keywords, listed in Table \ref{tab:keywords}.
	
	\begin{table}[h]
		\centering
		\caption{List of keywords identified by experts.}
		\begin{tabular}{l c}
			\hline
			Keyword& Type (Digit. Trans. or Shipping) \\
			\hline
			Digital transformation & Digit. Trans. \\
			Digital innovation & Digit. Trans. \\
			Digital ecosystems & Digit. Trans. \\
			Digitization & Digit. Trans. \\
			Digitalization & Digit. Trans. \\
			Digital platforms & Digit. Trans. \\
			Industry 4.0 & Digit. Trans. \\
			Smart technologies & Digit. Trans. \\
			Data-driven transformation & Digit. Trans. \\
			Automation & Digit. Trans. \\
			Internet of Things & Digit. Trans. \\
			Blockchain & Digit. Trans. \\
			Data analysis & Digit. Trans. \\
			Artificial intelligence & Digit. Trans. \\
			Machine learning & Digit. Trans. \\
			Big data & Digit. Trans. \\
			Cloud computing & Digit. Trans. \\
			Cyber-physical systems & Digit. Trans. \\
			Digital twins & Digit. Trans. \\
			Edge computing & Digit. Trans. \\
			5G networks & Digit. Trans. \\
			Predictive analytics & Digit. Trans. \\
			Cybersecurity & Digit. Trans. \\
			Supply chain integration & Digit. Trans. \\
			shipping & Shipping \\
			maritime & Shipping \\
			Sea freight & Shipping \\
			Smart ports & Shipping \\
			Autonomous ships & Shipping \\
			Fleet management & Shipping \\
			Cargo tracking & Shipping \\
			Digital shipyards & Shipping \\
			Port digitalization & Shipping \\
			Port automation & Shipping \\
			Vessel performance & Shipping \\
			\hline
		\end{tabular}
		\label{tab:keywords}
	\end{table}
	
	Data was collected from three research engines: EBSCO \citep{vaughan2011ebsco}, ProQuest \citep{cooke2017proquest}, and IEEE eXplore \citep{wilde2016ieee}. The search was performed on October the 22nd 2024. For each engine, we retrieved scientific articles containing any of the digital transformation related keywords and any of the shipping industry related keywords, in either their title or abstract. The exact query for each engine are available on request. We limited our results using the following criteria: \(a\) only English literature, and \(b\) only scientific contributions published in peer-reviewed journals. Table \ref{tab:searchres} shows the results.
	
	\begin{table}[h]
		\centering
		\caption{Number of retrieved articles per research engine.}
		\begin{tabular}{l c}
			\hline
			Engine & No. of scientific articles \\
			\hline
			EBSCO & 1904 \\
			ProQuest & 2011 \\
			IEEE eXplore & 300 \\
			\hline
		\end{tabular}
		\label{tab:searchres}
	\end{table}
	
	All search engines provided the digital object identifier for the articles. This allowed us to screen the resulting set and identify 2324 unique articles for the subsequent analysis. One challenge of using different data engines is the variety of attributes they return for each article. In order to have the same information for each article, we queried a fourth search engine for all the 2324 articles. We chose OpenAlex \citep{priem2022openalex}, which has been shown to be suitable for bibliometric analysis \citep{alperin2024analysis}. Our final result set comprised 2293 scientific publications.
	
	\subsection{Descriptive statistics}
	We started our analysis evaluating descriptive statistics across our article set. More specifically, we calculated:
	\begin{enumerate}
		\item the distribution of the number of publications per year;
		\item the distribution of publications across authors, identifying the most prolific authors;
		\item the distribution of publications across institutions, identifying the research centers with the highest number of publications;
		\item the distribution of publications across countries.
	\end{enumerate}
	
	\subsection{Co-authorship network analysis}
	As a second step, we built and analyzed the network of co-authorship. Network analysis was performed in Python, using the NetworkX package \citep{hagberg2008exploring}. We identified 7723 distinct authors. We built the network using authors as nodes, and setting bi-directional links between them if there existed at least one publication that they co-authored. For each link, we stored within the graph object information about the authors institutes and countries for further analysis.
	
	Figure XXX shows the degree distribution of the co-authorship network. The graph suggested a power-law distribution. To confirm on our assumption, we run statistical tests comparing the likelihood of power-law distribution against the exponential distribution, the log-normal distribution, and the truncated power-law distribution.
	
	Next, we focused on the largest connected component of the network, made of 883 authors and 2753 links between them. The choice of focusing on the largest component was dictated mostly by computational limitations.
	
	Working on the largest component, we applied the Louvein community \citep{blondel2008fast} algorithm to identify the major communities of authors and investigated the distribution of institutions and countries across communities (see Fig. [XXX]).
	
	To conclude, we analyzed the network for small-world behavior. More specifically, we calculated both the clustering coefficient and the average path length and compared them to random networks of equivalent size.
	
	\subsection{Co-citation network analysis}
	We built a co-citation network of nodes (i.e., articles) and links (i.e. co-citation between two articles). The resulting graph had 1298 nodes. The degree distribution was tested for power-law characteristics against other plausible distributions (exponential, log-normal, and truncated power-law).
	
	Next, we identified the most influential articles (i.e., the top 10 in terms of received citations). Our goal was to check if the most cited articles were literature reviews. As presented in the following section, this turned out not to be the case, allowing us to draw relevant considerations over the demand of SLRs at the conjunction of digital transformation and shipping industry.
	
	We then moved our attention to the top 20\% cited papers and analyzed their topics. To achieve this, we create a sub-network using only the top 20\% cited papers and applied the Louvein community algorithm \citep{blondel2008fast}. Next, for each community collected the titles and applied natural language processing (NLP) to model their topics (BERTTopic \citep{paulcombining}).
	
	To conclude, we applied different centrality measures to the top 20\% graph to identify the 5 most relevant articles. These were analyzed more in details in terms of covered research area, as a preliminary trend analysis, further developed in our next and last analysis section.
	
	\subsection{Thematic analysis}
	Working on the entire set of articles (2290) we performed a thematic analysis to identify the major topic of research. We pre-processed the titles with the following steps:
	\begin{enumerate}
		\item lemmatization to transform words into their root forms;
		\item removal of stop-words;
		\item removal of non alpha-numeric text.
	\end{enumerate}
	Next, we applied tokenization and embedded each title using BERT \citep{devlin2018bert}. The resulting vectors were analyzed for unsupervised clustering. More specifically, we adopted three method to identify the ideal number of clusters: the Elbow index \citep{cui2020introduction}, the Calinski-Harabasz index \citep{calinski1974dendrite}, and the Davies-Bouldin index \citep{davies1979cluster}.
	Having identified the best number of cluster, we applied the unsupervised K-means algorithm and calculated the centroid for each cluster.
	Next, we identified for each cluster the 10 articles closest to the corresponding centroid and applied BERTTopic to extract the common themes.
	
	We concluded our thematic analysis by building two word clouds. Using both titles and abstracts from all articles, we applied the TF-IDF algorithm to each word and use it as weight when building the clouds. The first cloud was built over the entire set of words in titles and abstracts, while the second cloud was built after removing all shipping related terms (hence focusing on the digital technologies only).
	
	\section{Results}
	In this section we present the results of our analysis. We then discuss them in the next section.
	
	\subsection{Descriptive statistics}
	Figure [XXX] shows the distribution of articles across years. Altough the first publications are dates as back as the 1960s, only from the year 2005 we witness an increasing interest in the effects of digital transformation within the shipping and maritime industry. The number of publication increased minimally and not steadily between 2005 and 2015. From 2015 onwards, we witness an exponential increase in the number of publications. After reaching a peak in 2023, the trend seem to have stabilized. Considering that our data was collected at the end of October 2024, we can reasonably argue that the year 2024 has not witnessed a signficant increase of publication, compared to the previous year.
	
	Figures [XXX]-[XXX] show the top 20 authors, the top 20 institutes, and the top 20 countries in terms of number of publications. Considering the authors, we note how the 0.03\% of all authors in our cohort (20 out of 7723) cover over 2.9\% of the total publications, suggesting a skewed distriution of publications across authors. When looking at the top insititutes, we see they cover over 21\% of the total publications (see Table \ref{tab:resdescinst}), while the top 5 countries cover up to 50\% of total publications (see Table \ref{tab:resdesccountry}). Looking deeper into the top insitute, one can notice how many of those Universityies have strong historical bindings with the sea. Consider, as examples, the Dalian Maritime University, the Shangai Maritime Univerity, and the Delft Technical University. Similarly, looking at the most rapresentative countries one can see they all have strong maritime industry and economy.
	
	\begin{table}[h]
		\centering
		\caption{Distribution of publications across institutions.}
		\begin{tabular}{r c c c}
			\hline
			Institution & No. of publications & Cumulative \% & \% Of total \\
			\hline
			Dalian Maritime University & 70 & 3.48 & 3.48\\
			Wuhan University of Technology & 44 & 5.67 & 2.19\\
			Norwegian University of Science and Technology & 44 & 7.86 & 2.19\\
			Shanghai Maritime University & 41 & 9.90 & 2.04\\
			Nanyang Technological University & 30 & 11.39 & 1.50\\
			University of Rijeka & 23 & 12.53 & 1.14\\
			Hong Kong Polytechnic University & 17 & 13.38 & 0.85\\
			University of Strathclyde & 16 & 14.17 & 0.80\\
			University of Piraeus & 15 & 14.92 & 0.75\\
			Liverpool John Moores University & 14 & 15.61 & 0.70\\
			Istanbul Technical University & 13 & 16.26 & 0.65\\
			Tsinghua University & 13 & 16.91 & 0.65\\
			Delft University of Technology & 12 & 17.50 & 0.60\\
			Universidad Politécnica de Madrid & 12 & 18.10 & 0.60\\
			Korea Maritime and Ocean University & 11 & 18.65 & 0.55\\
			Zhejiang Ocean University & 11 & 19.19 & 0.55\\
			University of Genoa & 11 & 19.74 & 0.55\\
			National Technical University of Athens & 11 & 20.29 & 0.55\\
			University of South-Eastern Norway & 10 & 20.79 & 0.50\\
			Aalto University & 10 & 21.28 & 0.50\\
			\hline
		\end{tabular}
		\label{tab:resdescinst}
	\end{table}

	\begin{table}[h]
		\centering
		\caption{Distribution of publications across countries.}
		\begin{tabular}{r c c c}
			\hline
			Country ISO & No. of publications & Cumulative \% & \% Of total \\
			\hline
			CN & 487 & 24.29 & 24.29\\
			US & 138 & 31.17 & 6.88\\
			GB & 107 & 36.51 & 5.34\\
			KR & 96 & 41.30 & 4.79\\
			NO & 95 & 46.03 & 4.73\\
			GR & 66 & 49.33 & 3.29\\
			IN & 59 & 52.27 & 2.94\\
			ES & 58 & 55.16 & 2.89\\
			IT & 54 & 57.86 & 2.69\\
			SG & 54 & 60.55 & 2.69\\
			DE & 50 & 63.04 & 2.49\\
			CA & 47 & 65.39 & 2.34\\
			ID & 42 & 67.48 & 2.09\\
			AU & 40 & 69.48 & 2.00\\
			HR & 39 & 71.42 & 1.95\\
			PL & 38 & 73.32 & 1.90\\
			TR & 36 & 75.11 & 1.80\\
			SE & 30 & 76.61 & 1.50\\
			FI & 29 & 78.05 & 1.45\\
			JP & 26 & 79.35 & 1.30\\
			\hline
		\end{tabular}
		\label{tab:resdesccountry}
	\end{table}

	\subsection{Co-authorship network analysis}
	
	\subsection{Co-citation network analysis}
	
	\subsection{Thematic analysis}
	
	\section{Discussion}
	
	\section{Conclusion}
	Summarize key findings and future work.
	
	\bibliographystyle{elsarticle-harv}
	\bibliography{dxshippingbibliography}
	
\end{document}  % End of document